### 2017-11-12 Draft2/Indexing.Growth.tex
Fix one thing, agonize over another

% Todo: There's an awful lot of lemmas. Make some into theorems.
\begin{lemma}
	The formulae for $\VarIndexBuffer$ and $\VarIndexElement$ are
	\begin{align*}
	\VarIndexBuffer &= \VarUseful|_{n = i + 1}\\
	\VarIndexElement &= i - \VarGrowSeq_{\VarIndexBuffer - 1}
	\end{align*}
\end{lemma}

@@

	To find $\VarIndexElement$, consider the equation:
	\begin{align*}
	n = \text{\# of items in tail} + \text{\# of items in head}
	\end{align*}
	From the premise, $n = i + 1$. The number of items in the tail is the last size at which $\FuncGrow$ was called, or $\VarGrowSeq_{\VarUseful - 1} = \VarGrowSeq_{\VarIndexBuffer - 1}$. Finally, since the desired item is the last item in the growth array, it is also the last item in the head. Then the size of the head is $\VarIndexElement + 1$. Substituting these values into the above equation, I receive
	\begin{align*}
	i + 1 &= \VarGrowSeq_{\VarIndexBuffer - 1} + \VarIndexElement + 1\\
	\VarIndexElement &= i - \VarGrowSeq_{\VarIndexBuffer - 1}
	\end{align*}
	completing the proof.
\end{proof}

% Todo: Mention formulae for \VarUseful and \VarGrowSeq_{\VarIndexBuffer - 1}.
Now, the $\FuncDecompose$ function can be implemented as follows:

% Todo: Continue to consider usage of mathematical vs. algorithm vars.
% Todo: Can ceiling of \log_2 be implemented fast? Also, what if \VarInitCapacity is not a power of 2?
\begin{algorithm}
	\begin{algorithmic}
		\Function{$\FuncDecompose$}{$\ParamIndex$}
			\State $\LclIndexBuffer \gets \max(\left\lceil \log_{\VarGrowthFactor} (i + 1) - \log_{\VarGrowthFactor} \VarInitCapacity \right\rceil, 0)$
			\State $\LclIndexElement \gets \ParamIndex - \VarGrowthFactor^{\LclIndexBuffer - 1} \times \VarInitCapacity$
			\State \Return $(\LclIndexBuffer,\ \LclIndexElement)$
		\EndFunction
	\end{algorithmic}
\end{algorithm}

Clearly, $\FuncDecompose$ runs in constant time. At first glance, however, it appears to be quite expensive since it % Bookmark

>>>>

% Todo: There's an awful lot of lemmas. Make some into theorems.
\begin{lemma}
	The formulae for $\VarIndexBuffer$ and $\VarIndexElement$ are
	\begin{align*}
	\VarIndexBuffer &= \VarUseful|_{n = i + 1}\\
	\VarIndexElement &= i - \VarGrowSeq_{\VarIndexBuffer - 1}
	\end{align*}
	(By convention, $\VarGrowSeq_{-1} = 0$.)
\end{lemma}

@@

	To find $\VarIndexElement$, consider the identity:
	\begin{align*}
	n = \text{\# of items in tail} + \text{\# of items in head}
	\end{align*}
	First, I will determine the number of items in the tail. In the case where $\VarIndexBuffer = 0$, then because $\VarNumItemsTail = \VarIndexBuffer$, the tail contains $0$ buffers and thus $0 = \VarGrowSeq_{-1}$ items. If $\VarIndexBuffer > 0$, the number of items in the tail is the last size at which $\FuncGrow$ was called, or the last term of $\VarGrowSeq$. This quantity is $\VarGrowSeq_{\VarUseful - 1} = \VarGrowSeq_{\VarIndexBuffer - 1}$.

	The desired item is the growth array's last element, which implies it is also the head's last element. Thus $\VarIndexElement$ is the head's last valid index, so the head size is $\VarIndexElement + 1$. Finally, from the premise, $n = i + 1$. Substituting all values into the above identity, I receive
	\begin{align*}
	i + 1 &= \VarGrowSeq_{\VarIndexBuffer - 1} + \VarIndexElement + 1\\
	\VarIndexElement &= i - \VarGrowSeq_{\VarIndexBuffer - 1}
	\end{align*}
	completing the proof.
\end{proof}

Using the formulae for $\VarUseful$ and $\VarGrowSeq_i$, I now implement the $\FuncDecompose$ function:

% Todo: Continue to consider usage of mathematical vs. algorithm vars.
% Todo: Can ceiling of \log_2 be implemented fast? Also, what if \VarInitCapacity is not a power of 2?
\begin{algorithm}
	\begin{algorithmic}
		\Function{$\FuncDecompose$}{$\ParamIndex$}
			\State $\LclIndexBuffer \gets \max(\left\lceil \log_{\VarGrowthFactor} (\ParamIndex + 1) - \log_{\VarGrowthFactor} \VarInitCapacity \right\rceil, 0)$
			\If{$\VarIndexBuffer > 0$}
				\State $\LclIndexElement \gets \ParamIndex - \VarGrowthFactor^{\LclIndexBuffer - 1} \times \VarInitCapacity$
			\Else
				\State $\VarIndexElement \gets \ParamIndex$
			\EndIf
			\State \Return $(\LclIndexBuffer,\ \LclIndexElement)$
		\EndFunction
	\end{algorithmic}
\end{algorithm}

Clearly, $\FuncDecompose$ runs in constant time. Despite that, it still appears to be quite expensive: normally, logarithms and exponentiation require use of costly floating-point instructions. However, in the special case where $\VarGrowthFactor = 2$ and $\VarInitCapacity = 2^\VarLogInitCapacity$ for some constant whole number $\VarLogInitCapacity$, I claim that $\VarIndexBuffer$ and $\VarIndexElement$ can be computed without use of floating-point instructions.

To see this, first note that $\VarIndexBuffer$ becomes $\max(\left\lceil \log_2 (i + 1) \right\rceil - \VarLogInitCapacity, 0)$. There is a constant time algorithm to compute $\left\lceil \log_2 k \right\rceil$ without using floating-point for any $k \in \mathbb{N}$, which I will not discuss in this paper due to length concerns. (An implementation of the algorithm may be found in [Implementations section].) Equipped with such an algorithm, it is easy to see that the whole expression can be computed without use of floating-point.

In calculating $\VarIndexElement$, the only potential use of floating-point instructions comes from $\VarGrowthFactor^{\VarIndexBuffer - 1}$. When $\VarGrowthFactor = 2$, however, this can be calculated with a simple bit shift. Thus, it is not necessary to use floating-point instructions to calculate either $\VarIndexBuffer$ or $\VarIndexElement$.
